A Critical Appraisal of Zero-Multiple Method in Simulation Biomolecular Systems

### 2014-06-18 writedown/zm/report/report.tex
save

\begin{table}
  \centering
  \caption{The excess chemical potential $\mu^\exc$, constant volume molar heat capacity $C_{v,m}$, dielectric constant $\eps$, diffusion constant $D$ and viscosity $\eta$
    calculated by different methods.    
    The parenthesises in the last column show the statistical uncertainty
    at the confidence level of 95~\%.
    The bold numbers indicating that its deviation from the SPME result is larger than the statistical uncertainty.
  }
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccc cccc}\hline\hline
    Method      &   $r_c$ [nm] &    $l$     &  $\mu^\exc$ [kJ/mol] &$C_{v,m}$ [J/(mol K)] & $\eps$ (Sys.~$\systemsb$) & $D$ [$10^{-9}\textrm{m}^2/\textrm{s}$] &  $\eta$ [$10^{-3}\textrm{Pa}\cdot\textrm{s}$] \\\hline
    SPME        &         1.9     &       --      &       $-26.08$ (0.24) & 72.21 (0.97)  & 97.6 (3.4)  & 5.86 (0.07)   & 0.315 (0.007)\\
    RF          &         1.2     & --            &       $-26.03$ (0.32) & 72.28 (0.81)  & \textbf{59.4} (1.3)  & \textbf{6.27} (0.19)   & \textbf{0.449} (0.020)\\
    ZM          &         1.2     &       1       &       $-26.22$ (0.30) & 71.75 (0.89)  & 95.5 (2.4)  & \textbf{5.53} (0.11)   & \textbf{0.346} (0.006)\\ 
    ZM          &         1.2     &       2       &       $-26.24$ (0.24) & 71.53 (0.88)  & 95.4 (2.5)  & 5.82 (0.27)   & 0.318 (0.012)\\ 
    ZM          &         1.2     &       3       &       $-{26.18}$ (0.20) & 71.58 (0.88)  & 96.0 (2.7)  & 5.79 (0.09)   & 0.321 (0.008)\\ 
    ZM          &         1.2     &       4       &       $-{26.01}$ (0.26) & 71.07 (0.84)  & 99.7 (2.2)  & 5.89 (0.09)   & 0.318 (0.013)\\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp3}
\end{table}

\subsubsection{Constant volume molar heat capacity}
The constant volume molar heat capacity is defined by:
\begin{align}
  C_{v,m} = \frac{1}{k_BT N} \langle (\mathcal H - \langle\mathcal H\rangle)^2 \rangle,
\end{align}
where $\mathcal H$ is the Hamiltonian of the system. Since constant
volume heat capacity is a extensive thermodynamics quantity, it is
normalized by the number molecules in the system.
It is calculated in system $\systemsb$.

@@

>>>>

\begin{table}
  \centering
  \caption{The excess chemical potential $\mu^\exc$, constant volume molar heat capacity $C_{v,m}$, dielectric constant $\eps$, diffusion constant $D$ and viscosity $\eta$
    calculated by different methods.    
    The parenthesises in the last column show the statistical uncertainty
    at the confidence level of 95~\%.
    The bold numbers indicating that its deviation from the SPME result is larger than the statistical uncertainty.
  }
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccc ccccc}\hline\hline
    Method      &   $r_c$ &    $l$      & $\mu^\exc$  &$C_{v,m}$ &  $\eps$ & $D$ &  $\eta$ \\
                &    [nm] &             &   [kJ/mol] &[J/(mol K)] & (Sys.~$\systemsb$) & [$10^{-9}\textrm{m}^2/\textrm{s}$] &  [$10^{-3}\textrm{Pa}\cdot\textrm{s}$] \\\hline
    % Method      &   $r_c$ [nm] &    $l$      & $\mu^\exc$ [kJ/mol] &$C_{v,m}$ [J/(mol K)] & $\eps$ (Sys.~$\systemsb$) & $D$ [$10^{-9}\textrm{m}^2/\textrm{s}$] &  $\eta$ [$10^{-3}\textrm{Pa}\cdot\textrm{s}$] \\\hline
    SPME        &         1.9     &       --      &      $-26.08$ (0.24) & 72.21 (0.97)  & 97.6 (3.4)  & 5.86 (0.07)   & 0.315 (0.007)\\
    RF          &         1.2     & --            &      $-26.03$ (0.32) & 72.28 (0.81)  & \textbf{59.4} (1.3)  & \textbf{6.27} (0.19)   & \textbf{0.449} (0.020)\\
    ZM          &         1.2     &       1       &      $-26.22$ (0.30) & 71.75 (0.89)  & 95.5 (2.4)  & \textbf{5.53} (0.11)   & \textbf{0.346} (0.006)\\ 
    ZM          &         1.2     &       2       &      $-26.24$ (0.24) & 71.53 (0.88)  & 95.4 (2.5)  & 5.82 (0.27)   & 0.318 (0.012)\\ 
    ZM          &         1.2     &       3       &      $-{26.18}$ (0.20) & 71.58 (0.88)  & 96.0 (2.7)  & 5.79 (0.09)   & 0.321 (0.008)\\ 
    ZM          &         1.2     &       4       &      $-{26.01}$ (0.26) & 71.07 (0.84)  & 99.7 (2.2)  & 5.89 (0.09)   & 0.318 (0.013)\\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp3}
\end{table}
    % SPME        &         1.9     &       --      & 1.67049 (0.00014)     $-26.08$ (0.24) & 72.21 (0.97)  & 97.6 (3.4)  & 5.86 (0.07)   & 0.315 (0.007)\\
    % RF          &         1.2     & --            & 1.66733 (0.00016)      $-26.03$ (0.32) & 72.28 (0.81)  & \textbf{59.4} (1.3)  & \textbf{6.27} (0.19)   & \textbf{0.449} (0.020)\\
    % ZM          &         1.2     &       1       & 1.67052 (0.00022)     $-26.22$ (0.30) & 71.75 (0.89)  & 95.5 (2.4)  & \textbf{5.53} (0.11)   & \textbf{0.346} (0.006)\\ 
    % ZM          &         1.2     &       2       & 1.67029 (0.00013)     $-26.24$ (0.24) & 71.53 (0.88)  & 95.4 (2.5)  & 5.82 (0.27)   & 0.318 (0.012)\\ 
    % ZM          &         1.2     &       3       & 1.67024 (0.00018)     $-{26.18}$ (0.20) & 71.58 (0.88)  & 96.0 (2.7)  & 5.79 (0.09)   & 0.321 (0.008)\\ 
    % ZM          &         1.2     &       4       & 1.67031 (0.00010)     $-{26.01}$ (0.26) & 71.07 (0.84)  & 99.7 (2.2)  & 5.89 (0.09)   & 0.318 (0.013)\\

\begin{table}
  \centering
  \caption{The constant pressure molar heat capacity $C_{v,m}$, isothermal compressibility $\kappa_T$ and
    thermal expansion coefficient $\alpha$
    calculated by different methods.    
    The parenthesises in the last column show the statistical uncertainty
    at the confidence level of 95~\%.
    The bold numbers indicating that its deviation from the SPME result is larger than the statistical uncertainty.
  }
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccc ccc}\hline\hline
    Method      &   $r_c$ [nm] &    $l$      & $\kappa_T$ [$10^{-10}\textrm{m}^2/\textrm{N}$] & $C_{p,m}$ [J/(mol K)] & $\alpha$ [$10^{-3}\textrm{K}^{-1}$]\\\hline
    SPME        &         1.9     &       --      & 5.941 (0.050)               &79.18 (0.94)           &1.026 (0.016)          \\
    RF          &         1.2     & --            & \textbf{6.508} (0.058)      &\textbf{82.34} (0.95)  &\textbf{1.149} (0.018) \\
    ZM          &         1.2     &       1       & 5.872 (0.050)               &78.91 (0.95)           &1.009 (0.018)          \\
    ZM          &         1.2     &       2       & 5.952 (0.051)               &79.70 (0.96)           &1.033 (0.016)          \\
    ZM          &         1.2     &       3       & 5.991 (0.050)               &79.49 (1.01)           &1.024 (0.019)          \\
    ZM          &         1.2     &       4       & 5.961 (0.054)               &79.46 (0.98)           &1.034 (0.017)          \\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp3a}
\end{table}

% \subsubsection{Hydrogen bonding}
% We count

\subsubsection{Constant volume molar heat capacity}
The constant volume molar heat capacity is defined by:
\begin{align}
  C_{v,m} = \frac{1}{k_BT^2 N} \langle (\mathcal H - \langle\mathcal H\rangle)^2 \rangle,
\end{align}
where $\mathcal H$ is the Hamiltonian of the system. Since constant
volume heat capacity is a extensive thermodynamics quantity, it is
normalized by the number molecules in the system.
It is calculated in system $\systemsb$.

\subsubsection{Constant pressure molar heat capacity}
The constant pressure molar heat capacity is defined by:
\begin{align}
  C_{p,m} = \frac{1}{k_BT^2 N} \langle ( H - \langle H\rangle)^2 \rangle,
\end{align}
where $ H$ is the enthalpy of the system, defined by $H = \mathcal H + PV$. Since constant
pressure heat capacity is a extensive thermodynamics quantity, it is
normalized by the number molecules in the system.
It is calculated by a NPT simulation of  system $\systemsb$.

\subsubsection{Isothermal heat capacity}
The isothermal heat capacity is defined by
\begin{align}
  \kappa_T = \frac{1}{k_BT} \frac{\langle (V - \langle V\rangle)^2 \rangle}{\langle V\rangle}.
\end{align}
It is calculated by a NPT simulation of  system $\systemsb$.

\subsubsection{Thermal expansion coefficient}
The thermal expansion coefficient is defined by
\begin{align}
  \alpha = \frac{1}{k_BT^2\langle V\rangle} \langle (H - \langle H\rangle)\cdot(V - \langle V\rangle) \rangle
\end{align}
It is calculated by a NPT simulation of  system $\systemsb$.

@@

\subsection{Simulation with alanine dipeptide}

Here we denote the full phase space by $\Omega$, and the phase space
variable by $\vect x$. We assume that 1, the trajectory $\vect x(t)$
is Markovian; 2, it is ergodic and as $t\rightarrow\infty$ the phase
space probability density $p(\vect x)$ converges to the unique
equilibrium probability density $\mu(\vect x)$ and 3, the trajectory
is reversible (detailed balance condition is satisfied). One can
easily write down a propagator for the probability density
\begin{equation}
\label{eq:propagator}  
p_{t+\tau}(\vect x) = \mathcal Q(\tau)\circ
p_t(\vect x),
\end{equation}
where the constant $\tau$ is called the lag time. Due to the
Markovianity of the system, the propagator satisfies the
Chapman-Kolmogorov equation:
\begin{equation}
  \label{eq:chapman}
  p_{t+k\tau}(\vect x) = \mathcal Q^k(\tau)\circ p_t(\vect x).
\end{equation}
Since the dynamics is reversible, the eigenvalues of $\mathcal
Q(\tau)$ are real-valued. We consider the first $m$ dominant
eigenvalues from the spectrum: $1 = \lambda_1 > \lambda_2 \geq
\lambda_3 \geq \cdots \geq \lambda_m$. For most of the biomolecule
systems, it is reasonable to assume a spectrum gap, i.e.~for a certain
$m$ $\lambda_m$ is much larger than the magnitude of the rest
eigenvalues. For the first $m$ leading eigenvalues one define the
corresponding timescales by
\begin{equation}
  \label{eq:timescale}
  t_i = -\frac{\tau}{\log \lambda_i}, \quad i= 2,\cdots, m
\end{equation}
Usually the first $m$ leading timescales can be identified with the
slowest conformational changes of the biomolecule. The spectrum gap is
considered as a timescale separation between the slow conformational
changes and the fast motion in the system, e.g.~bond vibration, angle
vibration, solvent motion, and other relatively fast conformational
changes. The eigenfunction of the first eigenvalue is nothing but the
equilibrium probability density $\mu(\vect x)$. We denote the
eigenfunctions corresponding to $\lambda_2, \cdot, \lambda_m$ by
$\phi_2(\vect x), \cdots, \phi_m(\vect x)$, then the original dynamics
of the system~\eqref{eq:chapman} is approximated by
\begin{equation}
  \label{eq:app-dynamics}
  p_{t+k\tau}(\vect x) = \mu(\vect x) + \sum_{i=2}^m e^{-\frac{k\tau}{t_i}} a_i \phi_i(\vect x) + \textrm{fast decaying dynamics}
\end{equation}
where $a_i$ are constants calculated by $a_i = \int \frac{1}{\mu(\vect
  x)} p_t(\vect x)\phi_i(\vect x) d\vect x$. This means that the
evolution of the phase space probability is a summation of the
time-independent equilibrium probability and a series of slow decaying
processes that are defined by the timescales $t_i$ and eigenfunctions
$\phi_i(\vect x)$. The eigenfunctions indicates how the conformational
change happens.
The slowest timescales and the corresponding conformational
changes are usually of the most biological interest.  Since the
slowest timescales determine the convergence speed of the system with
respect to the equilibrium probability distribution, they are the most
difficult part to sample and estimate in molecular dynamics
simulations.

In practice, it is not easy to directly calculate the eigenvalues and
eigenfunctions from molecular trajectories. Therefore, the continuous
trajectories are usually projected on to a finite number of sets, so
that the original dynamics is reduced to a Markov jump process that is
also called the Markov state model. We expect the Markov jump process
is a good approximate to the original dynamics in the sense that the
eigen pairs of the reduced process accurately reproduce those of the
original dynamics. How to do the discretized and the accuracy are
explained in detailed in Ref.~\cite{prinz2011markov}.

\begin{figure}
  \centering  
  \includegraphics[width=0.49\textwidth]{fig/result.alanine.amber/c-2.eps}
  \caption{(a) A schematic plot of the alanine dipeptide and the
    dihedral angles $\phi$ and $\psi$. (b) The free energy landscape on the
    $\phi$-$\psi$ dihedral angle space.}
  \label{fig:free-energy}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{fig/result.alanine.amber/fig-dih-dist-phi.eps}\\
  \includegraphics[width=0.40\textwidth]{fig/result.alanine.amber/fig-dih-dist-psi.eps}
  \caption{The probability distributions of the dihedral angles.}
  \label{fig:dih-dist}
\end{figure}

We simulate the alanine dipeptide with the Amber
ff-99SB-ILDN~\cite{lindorff2010improved} force field.  The dihedral
angles $\phi$ and $\psi$ (see Fig.~\ref{fig:free-energy}~(a)) are
widely accepted as collective variables that can describe the
conformational dynamics of the system.  The free energy landscape
on $\phi$-$\psi$ space is given in Fig.~\ref{fig:free-energy}~(b).  The projected
probability distributions on both dihedral angles
with error bars are presented in Fig.~\ref{fig:dih-dist}. It can be easily seen
from the distribution $P(\psi)$ and $P(\psi)$ that the ZM method with
$l=2$ and 3 reproduce the over-all probability of the main
conformations, i.e.~the $\alpha$-helix (right-handed), $\beta$-sheet and
the left-handed $\alpha$-helix with satisfactory accuracy.  For these
methods, deviations of $P(\phi)$ at $\phi = -75^\circ$ and $\phi =
-140^\circ$ are observed. They correspond to the double-peak structure
of the right-handed $\alpha$-helix and $\beta$-sheet, which means the
error happens at the sub-structure of the right-handed $\alpha$-helix
and $\beta$-sheet, which are of relatively smaller biological interest
comparing with the main conformations.

\begin{table}
  \centering
  \caption{The first three non-trivial timescales in ps computed at lag-time $\tau = 20$~ps.
  }
  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}c   ccc}\hline\hline
    Method      &        $t_2$ [ps]  & $t_3$ [ps] &  $t_4$ [ps] \\\hline
    SPME        &         1292  (148)     &          69.95  (0.76)  &         36.0 (4.7)    \\
    RF          & \textbf{1615} (155)     &  \textbf{68.48} (0.37)  &         34.1 (1.6)    \\
    ZM $l=1$    & \textbf{1607} (218)     &  \textbf{72.08} (0.50)  &         31.4 (1.9)    \\
    ZM $l=2$    &         1324  (110)     &          69.97  (0.58)  &         40.4 (2.8)    \\
    ZM $l=3$    &         1312  (157)     &          70.50  (0.72)  &         37.1 (4.3)    \\
    ZM $l=4$    & \textbf{1635} (156)     &          69.71  (0.54)  &         39.2 (3.6)    \\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp3}
\end{table}

% \begin{table}
%   \centering
%   \caption{The leading time scales in ps.
%   }
%   \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}ccc ccc}\hline\hline
%     Method      &   $r_c$ &    $l$      & $t_2$  & $t_3$ &  $t_4$ \\\hline
%     SPME        &         1.9     &       --      &         1292  (148)     &          69.95  (0.76)  &         36.0 (4.7)    \\
%     RF          &         1.2     & --            & \textbf{1615} (155)     &  \textbf{68.48} (0.37)  &         34.1 (1.6)    \\
%     ZM          &         1.2     &       1       & \textbf{1607} (218)     &  \textbf{72.08} (0.50)  &         31.4 (1.9)    \\
%     ZM          &         1.2     &       2       &         1324  (110)     &          69.97  (0.58)  &         40.4 (2.8)    \\
%     ZM          &         1.2     &       3       &         1312  (157)     &          70.50  (0.72)  &         37.1 (4.3)    \\
%     ZM          &         1.2     &       4       & \textbf{1635} (156)     &          69.71  (0.54)  &         39.2 (3.6)    \\
%     \hline\hline
%   \end{tabular*}
%   \label{tab:tmp3}
% \end{table}

\begin{figure}
  \centering
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-spme-eig-vec-2.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-rf-eig-vec-2.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm2-eig-vec-2.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-2.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm4-eig-vec-2.eps}\\
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-spme-eig-vec-3.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-rf-eig-vec-3.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm2-eig-vec-3.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-3.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm4-eig-vec-3.eps}\\
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-spme-eig-vec-4.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-rf-eig-vec-4.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-rf-eig-vec-2-comp.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-rf-eig-vec-3-comp.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-rf-eig-vec-4-comp.eps}\\
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm2-eig-vec-4.eps}
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-4.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-5.eps}\\
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-2-comp.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-3-comp.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-4-comp.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm3-eig-vec-5-comp.eps}\\
  \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm4-eig-vec-4.eps}\\
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.amber/fig-zm4-eig-vec-5.eps}\\
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-spme-eig-vec-2.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-rf-eig-vec-2.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-2.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm4-eig-vec-2.eps}\\
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-spme-eig-vec-3.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-rf-eig-vec-3.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-3.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm4-eig-vec-3.eps}\\
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-spme-eig-vec-4.eps}
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-rf-eig-vec-4.eps}
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-rf-eig-vec-2-comp.eps}
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-rf-eig-vec-3-comp.eps}
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-rf-eig-vec-4-comp.eps}\\
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-4.eps}
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-5.eps}\\
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-2-comp.eps}
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-3-comp.eps}
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-4-comp.eps}
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm3-eig-vec-5-comp.eps}\\
  % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm4-eig-vec-4.eps}\\
  % % \includegraphics[width=0.19\textwidth]{fig/result.alanine.charmm27/fig-zm4-eig-vec-5.eps}\\
  \caption{The eigenfunctions corresponding to the largest three
    non-trivial timescales, computed with lag-time 20~ps.}
  \label{fig:tmp5}
\end{figure}

The leading timescales and corresponding eigenfunctions are given in
Tab.~\label{tab:tmp3} and Fig.~\ref{fig:tmp5}. The ZM method with
order 2 and 3 presents best accuracy.
